---
title: "Journal 1"
#bibliography: references.bib
author: "Rob Salas"
output: 
  html_document:
    css: tweaks.css
    toc:  true
    toc_float: true
    number_sections: false
 
---



```{r, globalsettings, echo=FALSE, warning=FALSE, results='hide'}
library(knitr)

knitr::opts_chunk$set(echo = TRUE)
opts_chunk$set(tidy.opts=list(width.cutoff=100),tidy=TRUE, warning = FALSE, message = FALSE,comment = "#>", cache=TRUE, class.source=c("test"), class.output=c("test2"))
options(width = 100)
rgl::setupKnitr()



colorize <- function(x, color) {sprintf("<span style='color: %s;'>%s</span>", color, x) }

```

```{r klippy, echo=FALSE, include=TRUE}
#klippy::klippy(position = c('top', 'right'))
#klippy::klippy(color = 'darkred')
#klippy::klippy(tooltip_message = 'Click to copy', tooltip_success = 'Done')
```

Last compiled on `r format(Sys.time(), '%B, %Y')`

<br>

----

```{r include=FALSE}
fsave <- function(x, file, location = "./data/processed/", ...) {
    if (!dir.exists(location))
        dir.create(location)
    datename <- substr(gsub("[:-]", "", Sys.time()), 1, 8)
    totalname <- paste(location, datename, file, sep = "")
    print(paste("SAVED: ", totalname, sep = ""))
    save(x, file = totalname)
}

fpackage.check <- function(packages) {
  lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  })
}

packages <- c("haven", "tidyverse")
fpackage.check(packages)
```

**What is the conclusion of the authors about inferring about a party position using the mean of expert responses?**

Isn't right because experts don't evaluate positions correctly, the mean create biased estimations of the true positions. The median is better because is less sensitive to high dispersion.

**Pick an item that reflects a policy dimension you are intested in, from either the POPPA or CHES dataset. Find Dutch parties that have approximately equal mean-scores on that particular item; and investigate how the shape of expert placement distributions varies across these parties. Use histograms to illustrate.**



```{r}
pop <- read_dta("./data/expert_data_stata.dta")
# subset Dutch parties (country id=19)
pop <- pop[which(pop$country_id == 19), ]
# subset party and dimensions names(pop)
pop <- pop[, c(4, 6:21)]

pop %>% 
  select(party, immigration) %>% 
  group_by(party) %>% 
  summarize(mean = mean(immigration, na.rm = TRUE),
            std = sd(immigration, na.rm = TRUE)) %>% 
  arrange(desc(mean)) %>% 
  mutate(dif=lag(mean)-mean) %>% 
  print(n=100)

```


```{r warning=FALSE, message=FALSE}
df_pop <- pop %>% 
  select(party, pop_immigration = immigration) %>% 
  filter(party %in% c('PvdA', 'CU')) 
  
df_pop %>% 
  ggplot(aes(x=pop_immigration))+
  geom_histogram(color="black", fill="white", position = "identity") +
  facet_wrap(~party, scales = "free_y", nrow = 2)
```


**Based on the article and your own empirical insights, make an informed decision about how to aggregate the expert-level responses regarding this dimension to the party-level. Use this strategy to aggregate the responses and construct a dataframe (named: df) with rows reflecting the parties and column reflecting the aggregated scores on the picked dimension(s).**



```{r}

df_pop_agg <- pop %>% 
  group_by(party) %>% 
  summarize(pop_immigration = mean(immigration, na.rm = T))

dpes <- read_spss("./data/DPES2021 v1.0.sav")

df <- dpes %>% 
  select(S152, N76) %>% 
  group_by(N76) %>% 
  summarize(dpes_sentence_death = median(S152, na.rm = T)) %>% 
  mutate(party = as.character(as_factor(N76))) %>% 
  select(-N76)

merged <- df_pop_agg %>% full_join(df, by='party')
merged %>% print(n=100)
```

```{r}
fsave(merged, "positions_data.RData")
```




